---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hapiverse

<!-- badges: start -->
<!-- badges: end -->

The hapiverse (human API universe) is a metapackage whose goal is embrace a
collection of R packages that work together to provide programmatic access to
the European Bioinformatics Institute (EMBL-EBI) REST API services, namely those
related with human genetic variation and phenotypes. Currently, the hapiverse
provides easy installation and loading of two packages:

- [gwasrapidd](https://github.com/ramiromagno/gwasrapidd) that provides access to the [GWAS Catalog](https://www.ebi.ac.uk/gwas/)
- [quincunx](https://github.com/maialab/quincunx) that provides access to the [PGS Catalog](https://www.pgscatalog.org/)

## Installation

You can install the current version of hapiverse with:

``` r
# install.packages("remotes")
remotes::install_github("maialab/hapiverse")
```

After installing hapiverse, you also get [gwasrapidd](https://github.com/ramiromagno/gwasrapidd) and
[quincunx](https://github.com/maialab/quincunx) installed.

## Usage

To load and attach concomitantly [gwasrapidd](https://github.com/ramiromagno/gwasrapidd) and
[quincunx](https://github.com/maialab/quincunx) simply run:

```{r}
library(hapiverse)
```

## Namespace collisions

During the attaching of hapiverse, you will get notified of those functions that have clashing names. Behind the scenes we use the package [conflicted](https://github.com/r-lib/conflicted) to avoid misuse of functions with the same name. In these cases you will be required to choose one of the functions by using the double colon operator (`::`).

As an example, both gwasrapidd and quincunx provide a function named `get_traits()`. If you try to use it directly, R will throw an error and ask that you specify either `gwasrapidd::get_traits()` or `quincunx::get_traits()`.

Trying to use `get_traits()` will return an error because of the ambiguity arising from the two definitions:

```{r error=TRUE}
get_traits()
```
Note that you can use the function `hapiverse_conflicts()` at any time to revise the list of clashing function names:

```{r}
hapiverse_conflicts()
```

If you specify the package name with the double colon operator (`::`), then the ambiguity is resolved and execution proceeds normally. Using `get_traits()` from gwasrapidd:

```{r}
gwasrapidd::get_traits(efo_id = 'EFO_0005537')
```

Or using `get_traits()` provided by quincunx:

```{r}
quincunx::get_traits(efo_id = 'EFO_0005537')
```
